---
title: "wildfire database development"
author: "S. Earl"
format: gfm
---

# overview ---------------------------------------------------------------------

Documentation and general workflow overview for developing the wildfire
database to support the CRASS 2.0 study that will investigate relationships
between wildfire and aquatic biogeochemistry in the arid western United States.
The focus here is strictly on development of the wildfire database that stores
information such as the geometry of study catchments, discharge, and water
chemistry data. Other workflows for harvesting, for example covariates, are
detailed in other files within this repository. The workflow draws heavily on
the [firearea](https://srearl.gitlab.io/firearea/index.html) package that also
support this effort. Important is that this is not a reproducible workflow:
there is a time-sensitive component to some of the `firearea` queries,
especially those related to discharge, such that there could be different
returns based on the date of query, and this workflow features iterative steps,
such as querying data for USGS sites in individual sites but for multiple
states that is not shown, and some, particularly, database actions are run once
after initial table construction but not when data are added subsequently.

This workflow documents adding non-usgs data (catchments, pour points) to the
wildfire database. Locations include NEON aquatic (Sycamore Creek, AZ; Red
Butte, UT), Santa Barbara Coastal LTER (SBC LTER), and Sycamore Creek LTREB
(Dos S; Below Mesquite Wash) monitoring sites. The NEON Sycamore Creek site is
co-located with the Dos S Sycamore Creek LTREB site and thus the LTREB Sycamore
Creek site reflects both the LTREB and NEON programs.

Whereas the USGS catchments, pour points, and chemistry and discharge data
employ the nhdplusTools suite of tools, the non-USGS site catchments are
delineated with Streamstats.


# helpers ----------------------------------------------------------------------

```{r}
#| eval: FALSE

dcss_possibly <- purrr::possibly(
  .f        = firearea::delineate_catchment_streamstats,
  otherwise = NULL
)
```

# site metadata ----------------------------------------------------------------

```{r}
#| eval: FALSE

(
  non_usgs <- readr::read_csv("data/non_usgs_data.csv") |> 
    janitor::clean_names() |> 
    dplyr::filter(!grepl("SYCR_SS", site_name, ignore.case = TRUE)) |> 
    dplyr::mutate(
      usgs_site = dplyr::case_when(
        grepl("bell 3", site_name, ignore.case = TRUE) ~ "hanan_bell_3",
        grepl("bell 4", site_name, ignore.case = TRUE) ~ "hanan_bell_4",
        grepl("sycamore", site_name, ignore.case = TRUE) ~ "neon_sycr",
        grepl("butte", site_name, ignore.case = TRUE) ~ "neon_rebu",
        grepl("Teakettle", site_name, ignore.case = TRUE) ~ "neon_tecr",
        grepl("onofre", site_name, ignore.case = TRUE) ~ "sbc_lter_ono",
        grepl("gaviota", site_name, ignore.case = TRUE) ~ "sbc_lter_gav",
        grepl("hondo", site_name, ignore.case = TRUE) ~ "sbc_lter_hon",
        grepl("refugio", site_name, ignore.case = TRUE) ~ "sbc_lter_ref",
        grepl("tecolote", site_name, ignore.case = TRUE) ~ "sbc_lter_tec",
        grepl("bell", site_name, ignore.case = TRUE) ~ "sbc_lter_bel",
        grepl("devereaux", site_name, ignore.case = TRUE) ~ "sbc_lter_dev",
        grepl("pedro", site_name, ignore.case = TRUE) ~ "sbc_lter_ped",
        grepl("atascadero", site_name, ignore.case = TRUE) ~ "sbc_lter_ata",
        grepl("atascadero", site_name, ignore.case = TRUE) ~ "sbc_lter_ata",
        grepl("burro", site_name, ignore.case = TRUE) ~ "sbc_lter_bur",
        grepl("nook", site_name, ignore.case = TRUE) ~ "sbc_lter_noo",
        grepl("mission", site_name, ignore.case = TRUE) & !grepl("nook", site_name, ignore.case = TRUE) ~ "sbc_lter_mis",
        grepl("rattle", site_name, ignore.case = TRUE) ~ "sbc_lter_rat",
        grepl("bmw", site_name, ignore.case = TRUE) ~ "sycr_bmw",
        TRUE ~ NA
      )
    )
)
```

# delineations and pour points -------------------------------------------------

## california

### hanan bell sites

#### pour points

```{r}
#| eval: FALSE

hanan_bell_pour_points <- dplyr::bind_rows(
  sf::st_read("data/hanan_bell_3_pp.geojson") |> 
    dplyr::rename(usgs_site = Name) |> 
    dplyr::mutate(usgs_site = "hanan_bell_3"),
  sf::st_read("data/hanan_bell_4_pp.geojson") |> 
    dplyr::rename(usgs_site = Name) |> 
    dplyr::mutate(usgs_site = "hanan_bell_4")
) |> 
  dplyr::select(-Name1)
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_pour_points"),
  value     = hanan_bell_pour_points,
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```

#### catchments

```{r}
#| eval: FALSE

hanan_bell_catchments <- dplyr::bind_rows(
  sf::st_read("data/hanan_bell_3_ws.geojson") |> 
    dplyr::rename(usgs_site = Name) |> 
    dplyr::mutate(usgs_site = "hanan_bell_3"),
  sf::st_read("data/hanan_bell_4_ws.geojson") |> 
    dplyr::rename(usgs_site = Name) |> 
    dplyr::mutate(usgs_site = "hanan_bell_4")
) |> 
  dplyr::select(usgs_site, geometry)
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_catchments"),
  value     = hanan_bell_catchments,
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```

### sbc lter

```{r}
#| eval: FALSE

(
  non_usgs_ca <- non_usgs |> 
    dplyr::filter(grepl("sbc", data_source, ignore.case = TRUE))
)
```

```{r}
#| eval: FALSE

non_usgs_ca_ws <- split(
  x = non_usgs_ca,
  f = non_usgs_ca$site_name
) |>
  {\(df) purrr::map_df(.x = df, ~ dcss_possibly(longitude = .x$longitude, latitude = .x$latitude, state_code = "ca", location_identifier = .x$usgs_site))}()
```

Examine output to identify sites that were not delineated.

```{r}
#| eval: FALSE

non_usgs_ca_ws |>
  dplyr::full_join(
    non_usgs_ca,
    by = c("location_id" = "usgs_site")
  ) |> 
  # print(n = Inf)
  dplyr::filter(sf::st_is_empty(geometry)) |> 
  sf::st_drop_geometry()
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_catchments"),
  value     = non_usgs_ca_ws |> dplyr::rename(usgs_site = location_id),
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```

### rattlesnake creek

It is generally, though not necessarily the case, that one or more locations
among the California sites will not delineate. This seems unrelated to the
sites but rather that `Streamstats` on occasion does not find a solution. For
sites that were not delineated in the purrr workflow, delineate and add sites
to the database individually. The example here is Rattlesnake Creek, which
seems to fail often when run as part of a batch, but could be any site.

```{r}
#| eval: FALSE

sbc_lter_rat_ws <- firearea::delineate_catchment_streamstats(
  longitude           = non_usgs[grepl("rattle", non_usgs$site_name, ignore.case = TRUE), ]$longitude,
  latitude            = non_usgs[grepl("rattle", non_usgs$site_name, ignore.case = TRUE), ]$latitude,
  state_code          = "ca",
  location_identifier = "sbc_lter_rat"
)
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_catchments"),
  value     = sbc_lter_rat_ws |> dplyr::rename(usgs_site = location_id),
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```


## arizona: ss + bmw

```{r}
#| eval: FALSE

non_usgs_az <- non_usgs |> 
  dplyr::filter(grepl("syc", site_name, ignore.case = TRUE))
```

```{r}
#| eval: FALSE

non_usgs_az_ws <- split(
  x = non_usgs_az,
  f = non_usgs_az$site_name
) |>
  {\(df) purrr::map_df(.x = df, ~ dcss_possibly(longitude = .x$longitude, latitude = .x$latitude, state_code = "az", location_identifier = .x$usgs_site))}()
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_catchments"),
  value     = non_usgs_az_ws |> dplyr::rename(usgs_site = location_id),
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```

## red butte canyon: the other (not sycr) neon site

```{r}
#| eval: FALSE

neon_rebu <- firearea::delineate_catchment_streamstats(
  longitude           = non_usgs[grepl("red", non_usgs$site_name, ignore.case = TRUE), ]$longitude,
  latitude            = non_usgs[grepl("red", non_usgs$site_name, ignore.case = TRUE), ]$latitude,
  state_code          = "ut",
  location_identifier = "neon_rebu"
)
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_catchments"),
  value     = neon_rebu |> dplyr::rename(usgs_site = location_id),
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```

## pour points (except Hanan's Bell sites that are already uploaded)

```{r}
#| eval: FALSE

non_usgs_pour_points <- non_usgs |> 
  dplyr::filter(
    grepl("sbc", data_source, ignore.case = TRUE) | grepl("syc|red", site_name, ignore.case = TRUE)
  ) |>
  sf::st_as_sf(
    coords = c("longitude", "latitude"),
    crs    = 4326
  ) |> 
  dplyr::select(
    usgs_site,
    geometry
  )
```

```{r}
#| eval: FALSE

DBI::dbWriteTable(
  conn      = pg,
  name      = c("firearea", "non_usgs_pour_points"),
  value     = non_usgs_pour_points,
  overwrite = FALSE,
  append    = TRUE,
  row.names = FALSE
)
```


# database configurations ------------------------------------------------------

```{r}
#| eval: FALSE

# run once after table creation

## add primary key

DBI::dbExecute(
  conn      = pg,
  statement = "ALTER TABLE firearea.non_usgs_catchments ADD PRIMARY KEY (usgs_site) ;"
)

DBI::dbExecute(
  conn      = pg,
  statement = "ALTER TABLE firearea.non_usgs_pour_points ADD PRIMARY KEY (usgs_site) ;"
)

## add index on geometry

DBI::dbExecute(
  conn      = pg,
  statement = "CREATE INDEX non_usgs_catchments_geometry_idx ON firearea.non_usgs_catchments USING GIST (geometry) ;"
)

DBI::dbExecute(
  conn      = pg,
  statement = "CREATE INDEX non_usgs_pour_points_geometry_idx ON firearea.non_usgs_pour_points USING GIST (geometry) ;"
)

## designate geometry type and crs

DBI::dbExecute(
  conn      = pg,
  statement = "
  ALTER TABLE firearea.non_usgs_catchments
  ALTER COLUMN geometry
  TYPE geometry(geometry, 4326)
  USING ST_SetSRID(geometry, 4326)
  ;
  "
)

DBI::dbExecute(
  conn      = pg,
  statement = "
  ALTER TABLE firearea.non_usgs_pour_points
  ALTER COLUMN geometry
  TYPE geometry(geometry, 4326)
  USING ST_SetSRID(geometry, 4326)
  ;
  "
)

## fix ring self-intersection

DBI::dbExecute(
  conn      = pg,
  statement = "
  UPDATE firearea.non_usgs_catchments
  SET geometry = st_makevalid(geometry)
  WHERE NOT st_isvalid(geometry)
  ;
  "
)

```

# add non_usgs_catchments to ecoregion_catchments ------------------------------

```{r}
#| eval: FALSE

DBI::dbExecute(
  conn      = pg,
  statement = "
  INSERT INTO firearea.ecoregion_catchments (usgs_site, geometry)
  SELECT usgs_site, geometry FROM firearea.non_usgs_catchments
  WHERE usgs_site !~* 'ono|dev'
  ;
  "
)

```

# fix NEON site names ----------------------------------------------------------

The names assigned to the NEON sites to this point were based on a combination
of 'neon_' and the first two letters of the site name (e.g., Sycamore Creek ~
'neon_sycr'). However, this is confusing because it diverges from the NEON name
(neon_sycr ~ SYCA and neon_rebu ~ REDB). Here, we change the names associated
with the NEON sites to the NEON site name designations (but lowercase).

```{r}
#| eval: TRUE

DBI::dbWithTransaction(
  conn = pg,
  {
    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.non_usgs_catchments
      SET usgs_site = 'redb'
      WHERE usgs_site = 'neon_rebu'
      ;
      "
    )

    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.non_usgs_catchments
      SET usgs_site = 'syca'
      WHERE usgs_site = 'neon_sycr'
      ;
      "
    )

    ## non_usgs_pour_points

    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.non_usgs_pour_points
      SET usgs_site = 'redb'
      WHERE usgs_site = 'neon_rebu'
      ;
      "
    )

    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.non_usgs_pour_points
      SET usgs_site = 'syca'
      WHERE usgs_site = 'neon_sycr'
      ;
      "
    )

    ## ecoregion_catchments 

    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.ecoregion_catchments
      SET usgs_site = 'redb'
      WHERE usgs_site = 'neon_rebu'
      ;
      "
    )

    DBI::dbExecute(
      conn      = pg,
      statement = "
      UPDATE firearea.ecoregion_catchments
      SET usgs_site = 'syca'
      WHERE usgs_site = 'neon_sycr'
      ;
      "
    )
  }
)

```

# neon data --------------------------------------------------------------------

## discharge

```{r}
#| eval: TRUE

neon_discharge <- function(neon_site) {

  q <- neonUtilities::loadByProduct(
    dpID       = "DP4.00130.001",
    # startdate = "2016-01",
    # enddate   = "2022-03",
    site       = c(neon_site),
    package    = "basic",
    check.size = FALSE
  )

  q <- q$csd_continuousDischarge

  site_name <- tolower(neon_site)

  q_upload <- q |> 
    dplyr::mutate(date = lubridate::date(endDate)) |> 
    dplyr::summarise(
      mean = mean(
        x     = maxpostDischarge,
        na.rm = TRUE
      ),
      sd = sd(
        x     = maxpostDischarge,
        na.rm = TRUE
      ),
      missing = sum(is.na(maxpostDischarge)),
      .by  = date
    ) |> 
    dplyr::mutate(usgs_site = site_name) |> 
    dplyr::select(
      usgs_site,
      tidyselect::everything()
    ) |> 
    dplyr::filter(!is.na(mean))

    DBI::dbWriteTable(
      conn      = pg,
      name      = c("firearea", "non_usgs_discharge"),
      value     = q_upload,
      overwrite = FALSE,
      append    = TRUE,
      row.names = FALSE
    )

}

purrr::walk(
  .x = c("REDB", "SYCA"),
  .f = ~ neon_discharge(neon_site = .x)
)

DBI::dbExecute(
  conn      = pg,
  statement = "
  COMMENT ON COLUMN firearea.non_usgs_discharge.mean IS 'liters per second' ;"
)
```

## exo

```{r}
#| eval: TRUE

neon_exo <- function(neon_site) {

  exo <- neonUtilities::loadByProduct(
    dpID       = "DP1.20288.001",
    # startdate = "2016-01",
    # enddate   = "2022-03",
    site       = c(neon_site),
    package    = "basic",
    check.size = FALSE
  )

  exo_data <- exo$waq_instantaneous
  # exo_vars <- exo$variables_20288

  site_name <- tolower(neon_site)

  exo_upload <- exo_data |> 
    dplyr::mutate(date = lubridate::date(endDateTime)) |> 
    dplyr::select(
      date,
      specificConductance,      # microsiemensPerCentimeter
      dissolvedOxygen,          # milligramsPerLiter
      localDissolvedOxygenSat,
      pH,
      chlorophyll,              # microgramsPerLiter
      turbidity,                # formazinNephelometricUnit
      fDOM                      # quinineSulfateUnit
    ) |>
    tidyr::pivot_longer(
      cols      = dplyr::where(is.numeric),
      names_to  = "analyte",
      values_to = "concentration"
    ) |>
    dplyr::summarise(
      mean = mean(
        x     = concentration,
        na.rm = TRUE
      ),
      sd = sd(
        x     = concentration,
        na.rm = TRUE
      ),
      missing = sum(is.na(concentration)),
      .by  = c(date, analyte)
    ) |> 
    dplyr::mutate(
      usgs_site = site_name,
      unit      = dplyr::case_when(
        analyte == "specificConductance"      ~ "microsiemensPerCentimeter",
        analyte == "dissolvedOxygen"          ~ "milligramsPerLiter",
        analyte == "localDissolvedOxygenSat"  ~ "percent",
        analyte == "pH"                       ~ "pH" ,
        analyte == "chlorophyll"              ~ "microgramsPerLiter",
        analyte == "turbidity"                ~ "formazinNephelometricUnit",
        analyte == "fDOM"                     ~ "quinineSulfateUnit"
      ),
      source = "NEON_exo"
    ) |> 
    dplyr::select(
      usgs_site,
      dplyr::everything()
    ) |> 
    dplyr::filter(!is.na(mean))

    DBI::dbWriteTable(
      conn      = pg,
      name      = c("firearea", "non_usgs_water_chem"),
      value     = exo_upload,
      overwrite = FALSE,
      append    = TRUE,
      row.names = FALSE
    )

}

purrr::walk(
  .x = c("REDB", "SYCA"),
  .f = ~ neon_exo(neon_site = .x)
)

```


## nitrate

```{r}
#| eval: TRUE

neon_nitrate <- function(neon_site) {

  nitrate <- neonUtilities::loadByProduct(
    dpID      = "DP1.20033.001",
    # startdate = "2016-01",
    # enddate   = "2022-03",
    site      = c(neon_site),
    package    = "basic",
    check.size = FALSE
  )

  nitrate_data <- nitrate$NSW_15_minute
  # nitrate_vars <- nitrate$variables_20033

  site_name <- tolower(neon_site)

  nitrate_upload <- nitrate_data |> 
    dplyr::mutate(date = lubridate::date(endDateTime)) |> 
    dplyr::select(
      date,
      surfWaterNitrateMean # micromolesPerLiter
    ) |>
    dplyr::summarise(
      mean = mean(
        x     = surfWaterNitrateMean,
        na.rm = TRUE
      ),
      sd = sd(
        x     = surfWaterNitrateMean,
        na.rm = TRUE
      ),
      missing = sum(is.na(surfWaterNitrateMean)),
      .by  = c(date)
    ) |> 
    dplyr::mutate(
      usgs_site = site_name,
      analyte   = "nitrate",
      unit      = "micromolesPerLiter",
      source    = "NEON_nitrate"
    ) |> 
    dplyr::select(
      usgs_site,
      date,
      analyte,
      dplyr::everything()
    ) |> 
    dplyr::filter(!is.na(mean))

    DBI::dbWriteTable(
      conn      = pg,
      name      = c("firearea", "non_usgs_water_chem"),
      value     = nitrate_upload,
      overwrite = FALSE,
      append    = TRUE,
      row.names = FALSE
    )

}

purrr::walk(
  .x = c("REDB", "SYCA"),
  .f = ~ neon_nitrate(neon_site = .x)
)

```

## grabs

```{r}
#| eval: TRUE

neon_grabs <- function(neon_site) {

  grabs <- neonUtilities::loadByProduct(
    dpID      = "DP1.20093.001",
    # startdate = "2012-01",
    # enddate   = "2022-02",
    site      = c(neon_site),
    package    = "basic",
    check.size = FALSE
  )

  grabs_data  <- grabs$swc_externalLabDataByAnalyte
  # grabs_field <- grabs$swc_fieldData
  # grabs_vars  <- grabs$variables_20093
  # grabs_cats  <- grabs$categoricalCodes_20093

  site_name <- tolower(neon_site)

  grabs_upload <- grabs_data |> 
    dplyr::mutate(date = lubridate::date(collectDate)) |> 
    dplyr::summarise(
      mean = mean(
        x     = analyteConcentration,
        na.rm = TRUE
      ),
      sd = sd(
        x     = analyteConcentration,
        na.rm = TRUE
      ),
      missing = sum(is.na(analyteConcentration)),
      .by  = c(date, analyte)
    ) |> 
    dplyr::mutate(
      usgs_site = site_name,
      unit      = NA_character_,
      unit      = dplyr::case_when(
        grepl("TDS",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Fe",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("SO4",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("CO3",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("conductance", analyte, ignore.case = TRUE) ~ "microsiemensPerCentimeter",
        grepl("TDP",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TPN",         analyte, ignore.case = TRUE) ~ "microgramsPerLiter",
        grepl("Cl",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("DIC",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("NO3+NO2",     analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("ANC",         analyte, ignore.case = TRUE) ~ "milliequivalentsPerLiter",
        grepl("Ca",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("HCO3",        analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Br",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TSS",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TDN",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("NH4N",        analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Si",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Mg",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Na",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TSS",         analyte, ignore.case = TRUE) ~ "microgram",
        grepl("TP",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TN",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TOC",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("K",           analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Mn",          analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("NO2",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("TPC",         analyte, ignore.case = TRUE) ~ "microgramsPerLiter",
        grepl("DOC",         analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("F",           analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("Ortho",       analyte, ignore.case = TRUE) ~ "milligramsPerLiter",
        grepl("ph",          analyte, ignore.case = TRUE) ~ "pH",
        TRUE ~ unit
      ),
      source = "NEON_grabs"
    ) |> 
    dplyr::select(
      usgs_site,
      dplyr::everything()
    ) |> 
    dplyr::filter(!is.na(mean))

    DBI::dbWriteTable(
      conn      = pg,
      name      = c("firearea", "non_usgs_water_chem"),
      value     = grabs_upload,
      overwrite = FALSE,
      append    = TRUE,
      row.names = FALSE
    )

}

purrr::walk(
  .x = c("REDB", "SYCA"),
  .f = ~ neon_grabs(neon_site = .x)
)
```


# add foreign keys to non-usgs tables

```{r}
#| eval: TRUE

dbExecute(
  conn      = pg,
  statement = "
  ALTER TABLE firearea.non_usgs_discharge
  ADD CONSTRAINT non_usgs_discharge_fk_usgs_site
  FOREIGN KEY (usgs_site) REFERENCES firearea.non_usgs_catchments(usgs_site)
  ;"
)

dbExecute(
  conn      = pg,
  statement = "
  ALTER TABLE firearea.non_usgs_pour_points
  ADD CONSTRAINT non_usgs_pour_points_fk_usgs_site
  FOREIGN KEY (usgs_site) REFERENCES firearea.non_usgs_catchments(usgs_site)
  ;"
)

dbExecute(
  conn      = pg,
  statement = "
  ALTER TABLE firearea.non_usgs_water_chem
  ADD CONSTRAINT non_usgs_water_chem_fk_usgs_site
  FOREIGN KEY (usgs_site) REFERENCES firearea.non_usgs_catchments(usgs_site)
  ;"
)


```

# San Dimas discharge and nitrate

## data

```{r}
#| eval: TRUE

bell3 <- readr::read_csv("../data/bell3_streamflow_nitrate_obs_20220127.csv")
bell4 <- readr::read_csv("../data/bell4_streamflow_nitrate_obs_20220127.csv")

```

## catchment area

```{r}
#| eval: TRUE

bell_3_area_m2 <- sf::st_read(
  dsn   = pg,
  query = "
  SELECT *
  FROM firearea.non_usgs_catchments
  WHERE usgs_site ~~* '%bell_3%'
  ;"
) |> 
  sf::st_area() |> 
  units::drop_units()

bell_4_area_m2 <- sf::st_read(
  dsn   = pg,
  query = "
  SELECT *
  FROM firearea.non_usgs_catchments
  WHERE usgs_site ~~* '%bell_4%'
  ;"
) |> 
  sf::st_area() |> 
  units::drop_units()

```

## discharge

```{r}
#| eval: TRUE

bell3_q <- bell3 |> 
  dplyr::mutate(
    flow = streamflow_mm * 0.001 * bell_3_area_m2 * 1000 * (1 / 86400)
  ) |> 
  dplyr::filter(!is.na(flow)) |> 
  glue::glue_data_sql("
    INSERT INTO firearea.non_usgs_discharge (
      usgs_site,
      date,
      mean
    )
    values
    (
      'hanan_bell_3',
      { date },
      { flow }
    )
    ;
    ",
    .con = DBI::ANSI()
  )

DBI::dbWithTransaction(
  conn = pg,
  {
    purrr::walk(
      .x = bell3_q,
      .f = ~ DBI::dbExecute(
        statement = .x,
        conn      = pg
      )
    )
  }
)

bell4_q <- bell4 |> 
  dplyr::mutate(
    flow = streamflow_mm * 0.001 * bell_3_area_m2 * 1000 * (1 / 86400)
  ) |> 
  dplyr::filter(!is.na(flow)) |> 
  glue::glue_data_sql("
    INSERT INTO firearea.non_usgs_discharge (
      usgs_site,
      date,
      mean
    )
    values
    (
      'hanan_bell_4',
      { date },
      { flow }
    )
    ;
    ",
    .con = DBI::ANSI()
  )

DBI::dbWithTransaction(
  conn = pg,
  {
    purrr::walk(
      .x = bell4_q,
      .f = ~ DBI::dbExecute(
        statement = .x,
        conn      = pg
      )
    )
  }
)

```

## nitrate

```{r}
#| eval: TRUE

bell3_no3 <- bell3 |> 
  dplyr::filter(!is.na(nitrate)) |> 
  glue::glue_data_sql("
    INSERT INTO firearea.non_usgs_water_chem (
      usgs_site,
      date,
      analyte,
      mean,
      unit,
      source
    )
    values
    (
      'hanan_bell_3',
      { date },
      'nitrate',
      { nitrate },
      'milligramsPerLiter',
      'EJH'
    )
    ;
    ",
    .con = DBI::ANSI()
  )

DBI::dbWithTransaction(
  conn = pg,
  {
    purrr::walk(
      .x = bell3_no3,
      .f = ~ DBI::dbExecute(
        statement = .x,
        conn      = pg
      )
    )
  }
)

bell4_no3 <- bell4 |> 
  dplyr::filter(!is.na(nitrate)) |> 
  glue::glue_data_sql("
    INSERT INTO firearea.non_usgs_water_chem (
      usgs_site,
      date,
      analyte,
      mean,
      unit,
      source
    )
    values
    (
      'hanan_bell_4',
      { date },
      'nitrate',
      { nitrate },
      'milligramsPerLiter',
      'EJH'
    )
    ;
    ",
    .con = DBI::ANSI()
  )

DBI::dbWithTransaction(
  conn = pg,
  {
    purrr::walk(
      .x = bell4_no3,
      .f = ~ DBI::dbExecute(
        statement = .x,
        conn      = pg
      )
    )
  }
)

```
